Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-11 docs/proposals/NSF/ProjDesc/ProjDesc.tex
more additions

 \documentclass[12pt]{article}

%---------------------------------------------------------------------------------------------------%
%% PACKAGES
\usepackage{fullpage}
%\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
%\usepackage[justification=RaggedRight]{caption}
\usepackage{outlines,enumitem}
\usepackage{wrapfig, todonotes}
%---------------------------------------------------------------------------------------------------%

@@

%---------------------------------------------------------------------------------------------------%
%% TITLE
\begin{document}
\title{Project Description --
Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media}

% Original erosion title
%Mathematical models and analysis for erosion and dissolution in porous-media flow}
%\author{}
\date{}
\maketitle
%---------------------------------------------------------------------------------------------------%

@@

We propose to extend our preliminary results for pure advection so that
concentrations with finite Peclet numbers can be investigated.
Numerical methods to solve the advection-diffusion
equation~\eqref{eqn:advectionDiffusion} in a complex geometry poses
several challenges that we will address.  We will start by splitting the
advection and diffusion equations using a Strang splitting~\cite{}.
Numerical solutions of the advection equation will be formed with a
semi-Lagrangian method. These methods have the advantage that they ...?
The forced diffusion equation is more challenging to solve in a complex
geometry with an integral equation formulation. One particular challenge
is extending the forcing function from the complex geometry and into a
regular geometry, such as a square, so that a high-order volume integral
or Fourier method can be applied. Recently, groups have used radial
basis functions to form a sufficiently smooth extension~\cite{}
\todo[inline]{A-K, MC}. Alternatively, others have used the space-time
heat kernel~\cite{}, but this requires storing an entire time history of
the solution. We propose to develop a new method that uses the Laplace
transform.

>>>>

\documentclass[11pt]{article}

%---------------------------------------------------------------------------------------------------%
%% PACKAGES
%\usepackage{fullpage}
%\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
%\usepackage[justification=RaggedRight]{caption}
\usepackage{outlines,enumitem}
\usepackage{wrapfig, todonotes}
\usepackage{titlesec}
\usepackage{fancyhdr}
%---------------------------------------------------------------------------------------------------%

\addtolength{\oddsidemargin}{-0.75in}
\addtolength{\evensidemargin}{-0.75in}
\addtolength{\textwidth}{1.5in}
\addtolength{\topmargin}{-0.75in}
\addtolength{\textheight}{1.5in}
% For 11pt size

\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\pagestyle{fancy}
\lhead{\footnotesize Moore and Quaife}
\chead{\footnotesize Erosion, Transport, and  Dispersion in Granular and
Porous Media}
\rhead{\footnotesize \thepage}
\cfoot{}

@@

%---------------------------------------------------------------------------------------------------%
%% TITLE
\begin{document}
\begin{center}
{\large \bf Erosion, Transport, and Dispersion in Granular
and Porous Media} 
\end{center}
%\title{Project Description --
%Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media}

% Original erosion title
%Mathematical models and analysis for erosion and dissolution in porous-media flow}
%\author{}
%\date{}
%\maketitle
%---------------------------------------------------------------------------------------------------%

@@

We propose to extend our preliminary results by considering
concentrations with finite Peclet numbers. Numerical methods to solve
the advection-diffusion equation~\eqref{eqn:advectionDiffusion} in a
complex geometry poses several challenges that we will address.  We will
start by splitting the advection and diffusion equations with a Strang
splitting method~\cite{str1968}.  Numerical solutions of the advection
equation will be formed with a semi-Lagrangian method~\cite{rob1981} so
that it can be solved on an Eulerian grid without a restrictive CFL
condition. PI Quaife has experience with using semi-Lagrangian
methods for another type of complex viscous flow~\cite{kab-qua-bir2017}.
The diffusion equation is more challenging to solve in a complex
geometry with an integral equation formulation.  One approach is to use
the space-time heat kernel~\cite{}, but this requires storing an entire
time history of the solution.  Alternatively, the heat equation can be
discretized in time and the resulting Yukawa equation is solved at each
time step~\cite{kro-qua2010}. A particular challenge of this method is
extending the forcing function from the complex geometry and into a
regular geometry, such as a square, so that a high-order volume integral
or Fourier method can be applied.  Recently, groups have used radial
basis functions to form a sufficiently smooth
extension~\cite{fry-kro-tor2019, fry-leh-tor2019}.

We propose to develop a new method that uses the Laplace transform.
Upon applying a Strang splitting, the resulting diffusion equation is
\begin{alignat}{3}
  \pd{c}{t} &= \frac{1}{\Pe} \Delta c, \quad &&\xx \in \Omega, \: t>0, \\
  \pd{c}{\nn} &= 0, &&\xx \in \partial\Omega, \: t>0, \\
  c(\xx,0) &= c_0(\xx), &&\xx \in \Omega.
\end{alignat}
Then, the Laplace transform, $C(\xx,s) = \mathcal{L}[c](\xx,t)$,
satisfies
\begin{align}
  sC(\xx,s) - \frac{1}{\Pe} = c_0(\xx).
\end{align}
This PDE will be solved with high-order accuracy and efficiently by
using fast-multipole-accelerated integral equation methods of the
Beltrami equation~\cite{kro-qua2011}, and the concentration will be
recovered with a inverese Laplace transform~\cite{jos-war2006}.
